---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r global_options}
knitr::opts_chunk$set(fig.path='Figs/')
```

##Setting Working Directory

```{r}
setwd("C:/Users/trint/Desktop")
df = read.csv("activity.csv")
head(df)
```

```{r}
library(dplyr)
#Calculate the total number of steps taken per day
total = df %>% group_by (date) %>% summarise (total = sum(steps,na.rm =T))

#Difference between a histogram and a barplot,
hist(total$total, xlab="No of steps", ylab="Day", main="Histogram of the total number of steps taken each day")
```

```{r}
#Mean and median of the total number of steps taken per day
mean = total %>% summarise (mean = mean(total))
mean
```

```{r}
median = median(as.numeric(total$total))
median
```

#time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
ts_interval = df %>% group_by (interval) %>% summarise (avg = mean(steps,na.rm =T))
with(ts_interval, plot(interval, avg, type = "l"))
```
```{r}
ts_interval = as.data.frame(ts_interval)
max_step = max(ts_interval[,"avg"])
ts_interval %>% filter (avg == max_step)
```
```{r}
# total number of missing values in the dataset 
length(which(is.na(df$steps) == T))
```

#missing values in the datasetimputed with mean for that day
```{r}
total = df %>% group_by (date) %>% summarise (total = sum(steps,na.rm =T))
total = as.data.frame(total)
count = df %>% group_by (date) %>% tally ()
count = as.data.frame(count)
total = merge(total,count,by="date")
total$mean = ifelse(total$total ==0,0, total$total/total$n)
total = total[,c("date","mean")]
df = merge(df,total, by="date")
df2 = df %>% group_by(date) %>% mutate (steps = ifelse(is.na(steps), mean, steps))
df2 = as.data.frame(df2)
df2 = df2[,-c(4)]
head(df2)                                       
```

#histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
#Calculate the total number of steps taken per day
```{r}
total = df2 %>% group_by (date) %>% summarise (total = sum(steps,na.rm =T))
total
```
#Difference between a histogram and a barplot,
```{r}
hist(total$total, xlab="No of steps", ylab="Day", main="Histogram of the total number of steps taken each day")
```

```{r}
#Mean and median of the total number of steps taken per day
mean2 = total %>% summarise (mean = mean(total))
mean2
mean
```

```{r}
median2 = median(as.numeric(total$total))
median2
median
```

#no diff for mean and median

#Are there differences in activity patterns between weekdays and weekends?
```{r}
library(dplyr)
df2$date = as.Date(df2$date)
df2 = df2 %>% mutate(Indicator = ifelse(weekdays(df2$date)=="Saturday" | weekdays(df2$date)=="Sunday", "Weekend", "Weekday"))
library(dplyr)
ts = df2 %>% group_by(Indicator, interval) %>% summarize(total=sum(steps))
```


```{r}
library(lattice)
knitr::opts_chunk$set(fig.path='Figs/')

with(ts, xyplot(total ~ interval | Indicator, 
            type = "l",      
            main = "Total Number of Steps within Intervals by dayType",
            xlab = "Daily Intervals",
            ylab = "Average Number of Steps"))

```
